\chapter{Work Plan} \label{chap:workplan}

\section*{}

This chapter describes the tasks to be accomplished during the
following 5 months. Each task will be subsequently divided into
subtasks and provide an estimated time.

The calendatization of the tasks during the next 5 months will
be presented in the section~\ref{sec:schedule}.

\section{Tasks} \label{sec:tasks}

\begin{description}

\item[Literature Review] \hfill

In order to implement key features of this research work, several
concepts and algorithms need to be deeper analysed and studied.

\begin{itemize}
  \item Explore \emph{OpenCV} face detection algorithms.
  \item Explore \evm{} method for heart rate detection.
  \item Explore \emph{OpenCV} feature detection and tracking algorithms.
  \item Explore noise reduction to improve heart rate detection
        using the \evm{} method.
\end{itemize}

Estimated time: 4 weeks.

\pagebreak

\item[Implementation] \hfill

Implement an \emph{Android} application for detecting a person
heart rate using the smartphone's camera.

\begin{itemize}
  \item Integrate \emph{Android} and \emph{OpenCV} library.
  \item Implement face detection using \emph{OpenCV} library.
  \item Implement real-time \emph{Augmented Reality} using
        \evm{} method.
  \item Implement heart rate detection based on the values
        obtained by the \evm{} method.
  \item Implement feature detection and tracking using \emph{OpenCV} library.
  \item Implement noise reduction solutions explored.
  \item Improve application design and user experience.
\end{itemize}

Estimated time: 10 weeks.

\item[Evaluation] \hfill

Compare the heart rate detected to the one obtained from the
\emph{Philips}~\cite{Philips2013} application -- \emph{Vital Signs
Camera} -- and to the measurement of an heart rate monitor or pulse
oximeter in different settings.

Estimated time: 4 weeks.

\item[Thesis Writing] \hfill

Thesis writing, including a detailed description of the algorithms
and methods used, and the results obtained.

Estimated time: 6 weeks.

\end{description}

\pagebreak

\section{Schedule} \label{sec:schedule}

Figure~\ref{fig:gantt} is a Gantt diagram containing the schedule
of the previously defined tasks for the next 5 months.

Tasks \emph{Literature Review} and \emph{Implementation} overlap
because important parts of the implementation will be heavily
influenced by a few articles. And, \emph{Implementation} and
\emph{Evaluation} overlap because the evaluation phase may provide
further insight for improving the implemented application and
algorithms.
